---
title: "Week6b-Lab assignment"
author: "Helia Tehrani"
date: "2023-03-02"
output:
 prettydoc::html_pretty:
    theme: cayman
    highlight: github 
---

## R Markdown

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      Warning=FALSE,
                      fig.path = "../Output/")
                      
```
# Welcome
My week_6 R markdown. In this file you will find

-  **Plot of Chemical Data**
-  **Table of Averaged Values per Variable in Chemical Data**

# Load Libraries
Loading libraries that I need to make my beautiful graph and table
```{r}
knitr::opts_chunk$set(message = FALSE)
library(here)
library(tidyverse)
library(palmerpenguins)
library(lubridate)
library(devtools)
library(kableExtra)
library(memer)
```

# Chemical Data Plot
I picked the option to plot the chemical data using a facet-wrapped box plot. The figure details every single value of the many variables gathered and divided by seasons.
```{r LabHomework_plot}
ChemData <- read_csv(here("Week_4","Data","chemicaldata_maunalua.csv")) #read data in
ChemData_long <- ChemData %>% #name chem data
  drop_na() %>% #drop na's
  pivot_longer(cols = Temp_in:percent_sgd, #pivot long
               names_to = "Variables", #give name to new column with variable names
               values_to = "Values") %>% # give name to column with values
  select(Site, Season, Variables, Values) #select data I want
ggplot(ChemData_long, #dataframe used
       aes(x = Season, #x axis
           y = Values, #y axis
           color = Season))+ #color by season
  geom_boxplot()+ #box plot
  facet_wrap(~Variables, scales = "free")+ #facet wrap box plot
  theme_light()+ #light background
  labs(title = "Values of Different Varibales Across Two Seasons") #labels
```

# Chemical Data Table
I've provided a table of all the average values for each variable in this section.
```{r Homework Table}
ChemData_long %>% #use long chem data I made
  group_by(Variables) %>% # group by variable
  summarise(Average_values = mean(Values, na.rm = TRUE)) %>% #summarise all values for variables
  kbl(caption = "Averaged Values per Variable") %>% #add title to table
  kable_styling(full_width = FALSE) %>% #make table smaller
  row_spec(c(1,3,5,7), background = "Orange")# make it a kable table #add color to every other row
```