---
title: "week7a-Lab-assignment"
author: "Helia Tehrani"
date: "2023-03-07"
output:
 prettydoc::html_pretty:
    theme: cayman
    highlight: github 
---

## R Markdown

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      Warning=FALSE,
                      fig.path = "../Output/")
```
# Welcome
My week_7a R markdown. In this file you will find

- maps of Plastic pollution.

# Load Libraries
Loading libraries that I need to make my graph (map).
```{r}
knitr::opts_chunk$set(message = FALSE)
library(here)
library(tidyverse)
library(palmerpenguins)
library(lubridate)
library(devtools)
library(kableExtra)
library(memer)
library(maps)
library(mapproj)
library(mapdata)
```
readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-06-11/meteorites.csv")
```{r Bring in Data}
meteorites <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-06-11/meteorites.csv") %>% #read in data from Github
  select(fall, year, lat, long) %>% #select columns 
  filter(year > 1912, #filtered to get 100 years of data
         fall == "Fell") #wanted meteorites that fell
#Data from USA
usa<-map_data("usa") #used usa map
state <- map_data("state") #used states of usa map for map
```
# Map
I created a map of the United States showing meteorites that have fallen there between 1913 and 2013 using the meteorite data. This is what I got, and I'm happy with it. I was trying to filter out longitude and latitude in the meteorite data to make my picture bigger, but I was having trouble.
```{r Homework Map Lab 1, out.width="125%"}
ggplot()+
  geom_polygon(data = state, #used state to get state lines drawn on US
               aes(x = long, 
                   y = lat, 
                   group = group,
                   fill = region),  #used fill 
               color = "blue")+
   geom_point(data = meteorites, # add a point at all my sites 
             aes(x = long, 
                 y = lat))+
    guides(fill=FALSE)+ #removed guide of states
  coord_map(projection = "mercator",
            xlim = c(-125, -20), #set longitude
            ylim = c(25, 48))+ #set latitude
  theme_void()+ #only show US, not gridlines
  labs(title = "Sites of Fallen Meteorites in USA Over 100 Years (1913-2013)")
```
