---
title: "Intro to {stringr}"
author: "Helia Tehrani"
date: "2023-04-19"
output:
 prettydoc::html_pretty:
    theme: cayman
    highlight: github 
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE,
                      fig.path = "../output/")
```

##Libraries to load
```{r} 
#libararies that we need
library(here)
library(tidyverse)
library(tidytext)
library(wordcloud2)
library(janeaustenr)
```

### Introduction of string
- What is a string?
  - a character that is within quotations
```{r}
words<-"This is a string" # a string given a name
words_vector<-c("Apples", "Bananas","Oranges") # several strings within a vector
```

### Manipulation

```{r}
paste("High temp", "Low pH") 
paste("High temp", "Low pH", sep = "-") 
paste0("High temp", "Low pH") 
shapes <- c("Square", "Circle", "Triangle") 
paste("My favorite shape is a", shapes)  
two_cities <- c("best", "worst") 
paste("It was the", two_cities, "of times.")
```
### Manipulation: Indiividual characters
```{r}
shapes # shapes vector
str_length(shapes) # how many letters are in each word?
seq_data<-c("ATCCCGTC")
str_sub(seq_data, start = 2, end = 4) # extract the 2nd to 4th AA
str_sub(seq_data, start = 3, end = 3) <- "A" 
seq_data
str_dup(seq_data, times = c(2, 3)) # times is the number of times to duplicate each string
```

### Whitespace
- Let's say that you did not copy and paste your treatments into your column as you should have done during the first week of class. Now that you have certain words with extra white spaces, R believes that you have created a brand-new word. Here's how to handle that situation.
```{r}
badtreatments<-c("High", " High", "High ", "Low", "Low")
badtreatments 
str_trim(badtreatments) ## this eliminates any whitespace in the middle but leaves it on either side.
str_trim(badtreatments, side = "left") # this removes left
str_pad(badtreatments, 5, side = "right") # after the fifth character, add a white space to the right.
str_pad(badtreatments, 5, side = "right", pad = "1") # Add a 1 to the right side of the line after the fifth character, but only to the fifth place; if there are three letters, the fourth and fourth will be whitespaced.
```
### Locale Sensitive 
- Important, these will perform differently in different places in the world/with different languages. The default is English, but you can set the language setting.

```{r}
x<-"I love R!"
str_to_upper(x) # Make everything upper case
str_to_lower(x) # Make it lower case
str_to_title(x) # Use title case. (Capitalize each word's first letter)
```
### Pattern matching
-stringr} has functions to view, detect, locate, extract, match, replace, and split strings based on specific patterns.

```{r}
 data<-c("AAA", "TATA", "CTAG", "GCTT")# View a particular pattern in a string vector.
str_view(data, pattern = "A") # discover every string with an A
str_view(data, pattern = "G") # discover every string with a G
str_detect(data, pattern = "A") # detect a specific pattern aka A
str_detect(data, pattern = "AT") # which ones have AT
str_locate(data, pattern = "AT") # locate a pattern
```
### Regex: regular expressions
- Regular expressions come in a variety of forms:
* Metacharacters: Regular expressions that match a single character are the most basic type. The majority of characters, including all letters and numerals, are self-matching regular expressions. There are some special characters for a language like R that have reserved meaning; these characters are known as "Metacharacters." additional significance outside of human language. Extended Regular Expressions (EREs) use the following metacharacters:
. \ | ( ) [ { $ * + ?
* Sequences: as the name suggests refers to the sequences of characters which can match. We have shorthand versions (or anchors) for commonly used sequences in R:
* Quantifiers
* Character classes
* POSIX character classes (Portable Operating System Interface)
```{r}
vals<-c("a.b", "b.c","c.d") # set of strings
str_replace(vals, "\\.", " ") # string, pattern, replace all the "." with a space
```
### Metacharacter: Aside About the Functions
- Each function in {stringr} has two forms a basic form that searches for the first instance of a character and a *_all that searches for all instances. 
```{r}
vals<-c("a.b.c", "b.c.d","c.d.e") # has multiple .
str_replace(vals, "\\.", " ") # string, pattern, replace
str_replace_all(vals, "\\.", " ") #only replaces the first occurence when using string, pattern, replace, and str_replace.
```

### Sequences
- subset the vector to only keep strings with digits
```{r}
val2<-c("test 123", "test 456", "test")
str_subset(val2, "\\d")
```
### Character Class
- A list of characters surrounded in square brackets is referred to as a character class or character set. Only one of the several characters can be matched using character sets. For instance, the regex character class [aA] matches both upper- and lowercase letters of the alphabet.
```{r}
str_count(val2, "[aeiou]") 
str_count(val2, "[0-9]") # count any digit
```
### Quantifiers
- Example - find the phone numbers
-Create a regex to locate all strings including phone numbers. We are aware that there is a particular pattern (3 numbers, 3 numbers, 4 numbers, and they can be separated by either a "." or a "-"). Assume further that we are aware that the first number cannot be a 1.
```{r}
strings<-c("550-153-7578",
         "banana",
         "435.114.7586",
         "home: 672-442-6739")
phone <- "([2-9][0-9]{2})[- .]([0-9]{3})[- .]([0-9]{4})" # {2} means any 2nd one of the previous one, separator can be a [-.]
```

### Tidytext
```{r}
head(austen_books()) #explore
tail(austen_books())
original_books <- austen_books() %>% ## get all of Jane Austen's books
  group_by(book) %>%
  mutate(line = row_number(), # find every line
         chapter = cumsum(str_detect(text, regex("^chapter [\\divxlc]", # count the chapters (starts with the word chapter followed by a digit or roman numeral)
                                                 ignore_case = TRUE)))) %>% #ignore lower or uppercase
  ungroup() # ungroup it so we have a dataframe again
# don't try to view the entire thing... its >73000 lines...
head(original_books)
tidy_books <- original_books %>%
  unnest_tokens(output = word, input = text) # add a column named word, with the input as the text column
head(tidy_books) # there are now >725,000 rows. Don't view the entire thing!
cleaned_books <- tidy_books %>%
  anti_join(get_stopwords()) # dataframe without the stopwords
#see an example of all the stopwords
head(get_stopwords())
head(cleaned_books) # by = join_by(world)
cleaned_books %>%
  count(word, sort = TRUE)
```
```{r}
sent_word_counts <- tidy_books %>%
  inner_join(get_sentiments()) %>% # only keep pos or negative words
  count(word, sentiment, sort = TRUE) # count them
```

```{r}
sent_word_counts %>%
  filter(n > 150) %>% # take only if there are over 150 instances of it
  mutate(n = ifelse(sentiment == "negative", -n, n)) %>% # add a column where if the word is negative make the count negative
  mutate(word = reorder(word, n)) %>% # sort it so it gows from largest to smallest
  ggplot(aes(word, n, fill = sentiment)) +
  geom_col() +
  coord_flip() +
  labs(y = "Contribution to sentiment")
```
Make a Wordcloud
```{r}
words<-cleaned_books %>%
  count(word) %>% # count all the words
  arrange(desc(n))%>% # sort the words
  slice(1:100) #take the top 100
wordcloud2(words, shape = 'triangle', size=0.3) # make a wordcloud out of the top 100 words
```
#END
