---
title: "Intro to basic linear modeling"
author: "Helia Tehrani"
date: "2023-04-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE,
                      fig.path = "../Output/")
```

##Intro to basic linear modeling

#loading libraries
```{r}
library(tidyverse)
library(here)
library(palmerpenguins)
library(broom)
library(performance) 
library(modelsummary)
library(tidymodels)
```

Model the penguin dataset.

We've tidied and visualized the data and have our set of hypotheses that we want to test. Now we can start modeling...

# Anatomy of a Basic Linear Model
To run a simple linear model you use the following formula:

mod<-lm(y~x, data = df)

lm = linear model, y = dependent variable, x = independent variable(s), df = dataframe.

You read this as y is a function of x

Multiple regression
mod<-lm(y~x1 + x2, data = df)

## Model the Penguin Dataset
```{r}
Peng_mod<-lm(bill_length_mm ~ bill_depth_mm*species, data = penguins)
```

```{r}
check_model(Peng_mod) 
```

## View results in base R
```{r}
anova(Peng_mod)
```

## View results with broom
```{r}
coeffs<-tidy(Peng_mod) # just put tidy() around it
coeffs # Tidy coefficients
```
```{r}
results<-glance(Peng_mod) 
results# tidy r2, etc
```

-augment Your original data should be supplemented with residuals and anticipated values, and the model and data must be placed together.
```{r}
resid_fitted<-augment(Peng_mod)
resid_fitted 
```

## Results in Model Summary
```{r}
Peng_mod_noX<-lm(bill_length_mm ~ bill_depth_mm, data = penguins) # NEW MODEL
models<-list("Model with interaction" = Peng_mod, # make a list of models and name them
             "Model with no interaction" = Peng_mod_noX)
modelsummary(models, output = here("Week_12","output","table.docx")) 
```

```{r}
modelplot(models) +
    labs(x = 'Coefficients', 
         y = 'Term names') +
    scale_color_manual(values = wes_palette('Darjeeling1'))
```

## Many Models with purrr, dplyr, and broom
```{r}
 models<- penguins %>%
  ungroup() %>% # the penguin data are grouped so we need to ungroup them
  nest(-species) %>% # nest all the data by species
  mutate(fit = map(data, ~lm(bill_length_mm~body_mass_g, data = .)))
# . = take this data set
models
models$fit
``` 
 
 
```{r}
 results<-models %>%
   mutate(coeffs = map(fit, tidy), # look at the coefficients
   modelresults = map(fit, glance)) %>%  # R2 and others
   select(species, coeffs, modelresults) %>% # only keep the results
   unnest() # put it back in a dataframe and specify which columns to unnest
results
```
## Tidymodels
```{r}
lm_mod <- linear_reg() %>%
  set_engine("lm") %>%
  fit(bill_length_mm ~ bill_depth_mm*species, data = penguins) %>%
  tidy()
lm_mod
```
### Pipe to a plot
```{r}
lm_mod<-linear_reg() %>%
  set_engine("lm") %>%
  fit(bill_length_mm ~ bill_depth_mm*species, data = penguins) %>%
  tidy() %>%
  ggplot()+
    geom_point(aes(x = term, y = estimate))+
    geom_errorbar(aes(x = term, ymin = estimate-std.error,
                      ymax = estimate+std.error), width = 0.1 )+
  coord_flip()
lm_mod
```