---
title: "Goodplot_Badplot"
author: "Helia Tehrani"
date: "2023-03-30"
output: 
  html_document: 
    theme: united 
    toc: TRUE 
    toc_float: TRUE
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE,
                      message = FALSE, 
                      fig.path="../Output/")
```
##Bad Plot 
For this plot, I will use the mtcars dataset that is included in R. My goal is to create a plot that is difficult to interpret and misleading.
## Load Libraries 
```{r}
library(tidyverse)
library(here)
library(ggplot2)
```
#Data
```{r}
ggplot(mtcars, aes(x = hp, y = mpg)) +
  geom_point(aes(size = qsec), color = "red", alpha = 0.5) +
  labs(title = "Cars: Power vs. Mileage",
       subtitle = "Using quarter mile time to adjust point size",
       x = "Horsepower",
       y = "Miles per Gallon") +
  scale_size_continuous(name = "Quarter mile time",
                         range = c(1, 20),
                         breaks = seq(0, 20, by = 5),
                         labels = c("Fast", "Moderately Fast", "Moderate", "Moderately Slow", "Slow")) +
  theme_bw() +
  theme(panel.grid = element_blank(),
        axis.line = element_line(size = 2),
        axis.ticks = element_line(size = 1, color = "black"),
        plot.title = element_text(size = 30, face = "bold", hjust = 0.5),
        plot.subtitle = element_text(size = 20, hjust = 0.5),
        axis.title = element_text(size = 20),
        legend.title = element_text(size = 20),
        legend.text = element_text(size = 15),
        legend.key.size = unit(2, "cm"),
        panel.border = element_rect(size = 1, color = "black"),
        panel.background = element_rect(fill = "white"),
        plot.background = element_rect(fill = "gray95"))

